---
title: "data_manipulation with dplyr"
output: github_document
date: "2022-09-22"
---

## Load libraries

```{r}
library(tidyverse)
```

## Load data 
```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)

pups_data = read_csv("./data/FAS_pups.csv")
pups_data = janitor::clean_names(pups_data)

```

## Select variables

```{r}
select(litters_df, group: gd_of_birth)
```

## Select certain columns you'd like to remove:

```{r}
select(litters_df, -pups_survive)
```

## Renaming

```{r}
select(litters_df, GROUP = group, LiTter_NuMbEr = litter_number)
```

```{r}
rename(litters_df, GROUP = group, LiTtEr_NuMbEr = litter_number)
```

## Helper functions within select

```{r}
select(litters_df, starts_with("gd"))
select(litters_df, ends_with("weight"))
```

## Using `everything()`, which is handy for reorganizing columns without discarding anything
```{r}
select(litters_df, litter_number, pups_survive, everything())
```

## Filter rows 

*`gd_of_birth == 20`
*`pups_born_alive >= 2`
*`pups_survive != 4`
*`!(pups_survive == 4)`
`!((pups_survive == 4) & (gd_of_birth == 20))`
*`group %in% c("Con7", "Con8")`

## Drop NAs
*`drop_na(litters_data)`

## Mutate

```{r}
mutate(litters_df,
       wt_gain = gd18_weight - gd0_weight,
       group = str_to_lower(group))
      # wt_gain_kg = wt_gain * 2.2
```

## Re-arranging or sorting data
# Prints first 10 rows of dataset
```{r}
head(arrange(litters_df, group, pups_born_alive), 10)
head(arrange(litters_df, desc(group), pups_born_alive), 10)
```

## Piping %>%

```{r}
litters_df =
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)) %>%
  drop_na(wt_gain)
```


